quollr: An R Package for Visualizing 2-D Models from Nonlinear Dimension Reductions in High-Dimensional Space

### 2025-12-10 paper-quollr.Rmd
Updated quollr github link
Updated link plot section to reduce the number of pages
Removed unnecessary code

The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.

@@

<!--add each step separately-->
<!--add hexbin notation image-->

```{r, eval=FALSE}
#| label: code-illustration
# Code to draw illustration for notation
## hexagon binning to have regular hexagons
hb_obj_notation <- hex_binning(
  nldr_scaled_obj = scurve_umap_obj, 
  b1 = 9, 
  q = 0.1)

a1_temp <- hb_obj_notation$a1
a2_temp <- hb_obj_notation$a2
l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))

## Data set with all centroids
all_centroids_df_temp <- hb_obj_notation$centroids
hex_grid_temp <- hb_obj_notation$hex_poly

hex_grid_temp45 <- hex_grid_temp |> 
  filter(h == 45)

start_pt <- all_centroids_df_temp |> 
  filter(h == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = diff(scurve_umap_obj$lim2)/diff(scurve_umap_obj$lim1)) # x2max = r2

# To move the rectangle to ignore the overlap with the centroids
# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
rect_adj <- tibble(x1 = -0.03, x2 = 0.03)

a1 <- tibble(x = all_centroids_df_temp$c_x[4],
             xend = all_centroids_df_temp$c_x[5],
             y = all_centroids_df_temp$c_y[21],
             yend = all_centroids_df_temp$c_y[21],
             label = expression(a[1]))
a2 <- tibble(x = all_centroids_df_temp$c_x[25],
             xend = all_centroids_df_temp$c_x[25],
             y = all_centroids_df_temp$c_y[25],
             yend = all_centroids_df_temp$c_y[33],
             label = expression(a[2]))
l <- tibble(x = hex_grid_temp45$x[2],
            xend = hex_grid_temp45$x[3],
            y = hex_grid_temp45$y[2],
            yend = hex_grid_temp45$y[3],
            label = expression(l))

hex_param_vis <- ggplot() + 
    geom_polygon(data = hex_grid_temp, 
                        aes(x = x, 
                            y = y, 
                            group = h),
                 fill = "white", 
                 color = "#bdbdbd") +
    geom_point(data = all_centroids_df_temp, aes(
      x = c_x, 
      y = c_y), 
      color = "#31a354", size = 0.9) +
    geom_point(data = start_pt, aes(x = c_x, 
                                    y = c_y), 
               color = "black") + 
    geom_rect(data=d_rect, 
              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
              fill = "white", 
              color = "black", 
              alpha = 0, 
              linewidth = 0.7) +
    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2max - rect_adj$x2)) + 
    annotate("text", x=d_rect$x1min - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,
                     label = "(0,0)", 
             hjust=-0.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1max - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,
                     label = "(0,1)", 
             hjust=1.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1min - rect_adj$x1,
                     y=d_rect$x2max - rect_adj$x2,
                     label = expression(group("(",
                        list(0, y[2][max]),")")),
            hjust=-0.1, vjust=1.2) +
    geom_segment(data=d_rect, aes(
      x = x1min  - rect_adj$x1, # 0 - 0.03, 
      y = -0.31, 
      xend = x1max - rect_adj$x1, #1 - 0.03, 
      yend = -0.31), #-0.35),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
                 color = "black")+
    annotate("text", x=0.5, y=-0.36, 
             label = expression(r[1]), color = "black") +
    geom_segment(data=d_rect, aes(
      x = -0.25, 
      y = x2min - rect_adj$x2, #0 - 0.05, 
      xend = -0.25, 
      yend = x2max - rect_adj$x2), #r2 - 0.05),
      arrow = arrow(length = unit(0.03, "npc"),
                       ends = "both"), 
                 color = "black")+ 
    annotate("text", x=-0.3, y=0.4, 
             label = expression(r[2]), color = "black") +
    geom_segment(data = a1, aes(
      x = x, #-0.1 + 0.2087578, 
      y = y, #-0.15, 
      xend = xend, #-0.1 + 0.2087578*2, 
      yend = yend), #-0.15),
      arrow = arrow(length = unit(0.03, "npc"),
        ends = "both"), 
        color = "black")+ # a1 = 0.2087578
    annotate("text", 
             x=(a1$x+a1$xend)/2, 
             y=a1$y, 
             label = expression(a[1]), 
             color = "black",
             vjust = 1.2) +
    geom_segment(data = a2, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + # a2 = 0.1807896
    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
             label = expression(a[2]), 
             color = "black", hjust=-0.2) +
    annotate("text", x=-0.18, y=-0.24, 
      label = expression(group("(", list(s[1], s[2]), ")")),
      color = "black") +
  geom_segment(data = l, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + 
    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
             label = expression(l), 
             color = "black", hjust=-0.2) +
  coord_equal()

```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos="!ht", fig.alt="The components of the hexagon grid illustrating notation.", eval=FALSE}

hex_param_vis
```

>>>>

The package is available on CRAN, and the development version is available at [github.com/JayaniLakshika/quollr](https://github.com/JayaniLakshika/quollr).

@@
